Dish: Distribution-oblivious Shell Scripting

### 2020-02-29 pldi/draft.tex
Refine the first part of 4.2

\kk{I have to mention that all stateless and pure commands fall into
  the category of commands that have one input stream (and possibly
  several input static files) and one output stream (and no(?!?)
  output static files?). These commands are the bulk of
  data-processing in the shell, and these are the ones that we handle
  in our system.}

\kk{Move this assumption in the transformation}
Nodes may read from more files than just the files of their
input stream (resp. write to more files than their output stream).
However, these files are static (\eg used for configuration or
logging) and do not represent streams of data, and therefore are not
considered part of the dataflow graph.
For now, we assume that these static
files can only be accessed by one node, thus ensuring that they do not
interfere with the distributed implementation.
%% Because of the assumption in section Front End \ref{}, that each file
%% exists only once in the dataflow graph, we can safely assume that
%% reading and writing to these static files does not interfere with the
%% distributed implementation, and doesn't alter the behaviour of the
%% program.

Before describing the different types of parallelization
optimizations, we formalize the intuition about stateless and pure
commands that was described earlier~\sx{cmd}.
%% \nv{TODOs
%%   (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me)
%% }
%

As mentioned there, stateless commands such as \ttt{tr} operate independently on
individual elements of the stream (characters, lines, or files)
without maintaining any state.
% Some examples include \ttt{grep}, \ttt{basename}, and \ttt{tr}.
Formally, a command $f$ is stateless if
it is ``linear'' with respect to concatenation, \ie 
% the following equation holds:
satisfying the following equation:

\[
\forall x, x', f(x.x') = f(x).f(x')
\]

@@

\[
\forall x, x', f(x.x') = r(m(x).m(x'))
\]

>>>>

\kk{Move this assumption in the transformation: Nodes may read from
  more files than just the files of their input stream (resp. write to
  more files than their output stream).  However, these files are
  static (\eg used for configuration or logging) and do not represent
  streams of data, and therefore are not considered part of the
  dataflow graph.  For now, we assume that these static files can only
  be accessed by one node, thus ensuring that they do not interfere
  with the distributed implementation.
%% Because of the assumption in section Front End \ref{}, that each file
%% exists only once in the dataflow graph, we can safely assume that
%% reading and writing to these static files does not interfere with the
%% distributed implementation, and doesn't alter the behaviour of the
%% program.
}

Having a dataflow graph representation allows us to define
parallelization optimizations as graph transformations. Before
describing the different types of optimizations, we formalize the
intuition about stateless and pure commands that was described
earlier~\sx{cmd}. We focus on parallelizing commands that produce one
output and consume their inputs one at a time, with the possible
exception of static input files that are used for configuration. More
formally, we parallelize commands that can be represented with
functions of type $f : D\kstar \times [D\kstar] \to D\kstar$, where
the first argument represents the concatenation of the inputs that are
consumed one at a time and the second argument represents the static
inputs. Most stateless and pure commands fall in this category
\kk{Maybe have an argument here that we show in the evaluation that
  handling these commands is adequate for a big variety of data
  processing scripts, from several microbenchmarks to some
  large-realistic macrobenchmarks}.
%% \nv{TODOs
%%   (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me)
%% }

As mentioned in \Cref{cmd}, stateless commands such as \ttt{tr} operate
independently on individual elements of the stream (characters, lines,
or files) without maintaining any state. In this work, we consider
lines as the data quantum, so commands that are file-stateless are not
considered stateless. Formally, a command $f$ is stateless if it is
``linear'' with respect to concatenation, \ie satisfying the following
equation:

\[
\forall x, x', s, f(x.x', s) = f(x, s).f(x', s)
\]

@@

\[
\forall x, x', s, f(x.x', s) = r(m(x, s).m(x', s), s)
\]
